On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-03 tt_intro.tex

\begin{figure}
  \begin{center}
    \begin{tabular}{ccc}

@@

At delay 1, a bead cannot collaborate with its successors in order to stabilize itself. 
In fact, there are just two ways for a bead to get stabilized at delay 1 (or the bead has no place to be stabilized around so that the system halts), as observed in \cite{DHOPRSST2018}. 
One is to be bound and the other is through a 1-in-1-out structure called the tunnel section. 
See Fig.~\ref{fig:TTT_tunnel}. 
A \textit{tunnel section} consists of four beads that occupy four neighbors of one free point. 
In order for an oritatam system to stabilize the bead $w[i]$ at the central point $p$ of a tunnel section, its predecessor $w[i-1]$ must be put at one of the two free neighbors of $p$. 
Thus, at the stabilization of $w[i]$, only one neighbor of $p$ is left free so that the successor $w[i+1]$ is to be stabilized there, even without being bound. 
In this case, the point where $w[i-1]$ is stabilized is considered to be an \textit{entrance of the tunnel} and the point where $w[i+1]$ is stabilized is considered as an \textit{exit of the tunnel}.

The behavior of an oritatami system at delay 1 can be described by a sequence of $S$ of $b$ (bound), $t_s$ (straight tunnel section), $t_o$ (obtuse-turn tunnel section), and $t_a$ (acute-turn tunnel section); priority is given to tunnel, that is, $S[i]$ is $t_s$ (resp.~$t_o$, $t_a$) if the $i$-th bead of the system is stabilized not only by being bonded but also through a straight (resp. obtuse-turn, acute-turn) tunnel section. 
We let $S$ take the value $\blacksquare$ for halt (due to the lack of free neighbors).

%Assume that four of the six neighbors of a point $p$ are occupied by beads $a_{j_1},a_{j_2},a_{j_3},a_{j_4}$ while the other two are free. We call such a point $p$ the $inside\ of\ a\ tunnel$ and points $p^\prime$ the $entrance\ of\ a\ tunnel$ except when $p^\prime$ is inside of a tunnel. If the beads $w[i-2]$ and $w[i-1]$ are stabilized respectively at one of the two free neighbors and at $p$ one after another, then the next bead $w[i]$ cannot help but be stabilized at the other free neighbor. In this way, $w[i]$ can get stabilized without being bound.

@@

\begin{definition}[Binding Capability]
Let $\alpha$ be an arity and $C = (P,w,H)$ be a conformation of arity at most $\alpha$.
Let $H_k = H \cap \{ (i,j) \mid \mbox{$i=k$ or $j=k$} \}$. 
Moreover, let $R_k$ be a set of neighbors of the point $P[k]$ that are free and reachable from $C$.
The binding capability of $C$ at arity $\alpha$, denoted by $\#bc_\alpha(C)$, is defined by $\sum^{|w|}_{k=1} \min \{\alpha-|H_k|, |R_k|\}$.
Whenever the arity is clear from context, we omit the subscript $\alpha$. 
\end{definition}

>>>>

\begin{figure}[tb]
  \begin{center}
    \begin{tabular}{ccc}

@@

At delay 1, a bead cannot collaborate with its successors in order to stabilize itself. 
In fact, there are just two ways for a bead to get stabilized at delay 1 (or the bead has no place to be stabilized around so that the system halts), as observed in \cite{DHOPRSST2018}. 
One is to be bound and the other is through a 1-in-1-out structure called the tunnel section. 
See Fig.~\ref{fig:TTT_tunnel}. 
A \textit{tunnel section} consists of four beads that occupy four neighbors of one free point. 
In order for an oritatam system to stabilize the bead $w[i]$ at the central point $p$ of a tunnel section, its predecessor $w[i-1]$ must be put at one of the two free neighbors of $p$. 
Thus, at the stabilization of $w[i]$, only one neighbor of $p$ is left free so that the successor $w[i+1]$ is to be stabilized there, even without being bound. 
In this case, the point where $w[i-1]$ is stabilized is considered to be an \textit{entrance of the tunnel section} and the point where $w[i+1]$ is stabilized is considered as its \textit{exit}.

The behavior of an oritatami system at delay 1 can be described by a sequence of $S$ of $b$ (bound), $t_s$ (straight tunnel section), $t_o$ (obtuse-turn tunnel section), and $t_a$ (acute-turn tunnel section); priority is given to tunnel, that is, $S[i]$ is $t_s$ (resp.~$t_o$, $t_a$) if the $i$-th bead of the system is stabilized not only by being bonded but also through a straight (resp. obtuse-turn, acute-turn) tunnel section. 
Let us introduce $t$ as a wildcard for $t_s$, $t_o$, and $t_a$. 
We also let $S$ take the value $\blacksquare$ for halt (due to the lack of free neighbors). 
A tunnel is a set of

@@

\begin{definition}[Binding Capability]
Let $\alpha$ be an arity and $C = (P,w,H)$ be a conformation of arity at most $\alpha$.
Let $H_k = H \cap \{ (i,j) \mid \mbox{$i=k$ or $j=k$} \}$. 
Moreover, let $R_k$ be a set of neighbors of the point $P[k]$ that are free and reachable from $C$.
The binding capability of $C$ at arity $\alpha$, denoted by $\#bc_\alpha(C)$, is defined by $\sum^{|w|}_{k=1} \min \{\alpha-|H_k|, |R_k|\}$.
The subscript $\alpha$ is omitted whenever it is clear from context. 
\end{definition}

We now prove that in ``almost all'' tunnels is a troll domiciled and robs the transcript of binding capability (the original story is from \cite{Pratchett1992}). 
Originally, we tried to find a troll in every tunnel but failed; a troll seems to dislike the very first bead $w[1]$, or its property that only $\alpha$+1 beads around may take all hands of $w[1]$ thanks to the absence of its predecessor; any other bead must be surrounded by at least $\alpha+2$ beads in order to be free from free hand because a bead cannot bind with its predecessor or successor. 
We call the bead $w[1]$ \textit{singular}. 
A tunnel is \textit{singular} if its entrance or exit is next to $w[1]$. 
There can be at most 3 tunnels around one bead so that no more than 3 tunnels can be singular. 
A singular tunnel will be denoted with the superscript $\times$ like $t_s^\times$ or $t^\times$. 
In contrast, the notation without $\times$ such as $t_s$ and $t$ shall imply their non-singularity.
